---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
library(tidyverse)
library(dplyr)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`.

```{r}
deaths <- av |>
  pivot_longer(cols = starts_with("Death"), names_to = "Time", values_to = "Death") |>
  mutate(Time = parse_number(Time)) 

```

Similarly, deal with the returns of characters.

```{r}
returns <- av |>
  pivot_longer(cols = starts_with("Return"), names_to = "Time", values_to = "Return") |>
  mutate(Time = parse_number(Time))
```

Based on these datasets calculate the average number of deaths an Avenger suffers.

```{r}
avg_deaths = length(which(deaths$Death == "YES")) / length(unique(deaths$URL))
print(avg_deaths)
```

## Individually

For each team member, copy this part of the report.

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.



### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

### Nadia's FiveThirtyEight Statement

"There's ... a 50 percent chance [Avengers] recovered from a second or third death."

#### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

```{r}
deaths2_3 <- av %>% select(Name.Alias, Death2, Return2, Death3, Return3) %>%
  filter(Death2 == "YES")

# 16 Avengers died a second time
died2 <- count(deaths2_3)
# 8 Avengers returned from their second death
returned2 <- deaths2_3 %>% filter(Return2 == "YES") %>% count()
# 2 Avengers died a third time
died3 <- deaths2_3 %>% filter(Death3 == "YES") %>% count()
# 1 Avenger returned from their third death
returned3 <- deaths2_3 %>% filter(Return3 == "YES") %>% count()

# Avengers returned from exactly half of their second or third deaths
(returned2+returned3) / (died2 + died3)
```

#### Include your answer

Of all Avengers in the dataset, 16 died at least twice with 8 returning from their second death, and 2 died at least three times with only 1 recovering from their third death. This results in a recovery rate of 50%, which matches the number in the FiveThirtyEight analysis.

### Deo Shaji

-   My statement to verify: Out of the 173 listed Avengers , my analysis found that 69 had died at least once since they had joined the avengers.

```{r}
total_deaths <- deaths |>
  filter(Death == 'YES') 


length(unique(total_deaths$URL))
```

-   I filtered the deaths dataset based on the Death column having a 'YES'. Then I filtered based on the unique URLs which gave me the number of avengers who had died atleast once since they had joined the team. My value of 69 matches with what the article originally mentioned. Initially, I was going to count the number of unique name.Alias but I noticed some of the rows were empty for name.Alias., which is why I used URL

## Grace Wu

#### FiveThirtyEight Statement
- "I counted 89 total deaths — some unlucky Avengers7 are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback."

#### Code
```{r}
length(which(deaths$Death == "YES"))
length(which(returns$Return == "YES"))
```
#### Answer

- I counted the total number of deaths by counting the number of times Death in the deaths data set returned "YES". Then I counted the total number of returns by counting the number of times Return in the data set returned "YES". My value for total deaths was 89, and my value for total returns was 57, which aligns with the analysis statement above.

## Jenna Whalen
```{r}
total_comebacks <- returns %>%
  filter(Return == "YES") %>%
  summarise(comebacks = n())

total_comebacks
```
- I filtered the Return column to only include those that had returned from the dead ('YES'). I then used the 'summarise' function to verify there were 57 cases in which someone returned from the dead.
